---
title: "Does the Brain’s E/I Balance Really Shape Long-Range Temporal Correlations?"
titlerunning: ""
authors:
  - name: Lydia Sochan
    orcid: 
    email: lydiasochan@gmail.com
    roles:
      - data curation
      - formal analysis
      - investigation
      - software
      - visualization
      - writing - original draft preparation
    affiliations:
      - ref: 1
    degrees:
      - BMSc
      - MASc
  - name: Alexander Mark Weber
    orcid: 0000-0001-7295-0775
    corresponding: true
    email: aweber@bcchr.ca
    roles:
      - conceptualization
      - data curation
      - formal analysis
      - funding acquisition
      - investigation
      - methodology
      - project administration
      - resources
      - supervision
      - validation
      - visualization
      - writing - review & editing
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
    degrees:
      - PhD
      - MSc
affiliations:
  - id: 1
    name: School of Biomedical Engineering, The University of British Columbia, Vancouver, BC, Canada
  - id: 2
    name: BC Children's Hospital Research Institute, The University of British Columbia, Vancouver, BC, Canada
  - id: 3
    name: Pediatrics, The University of British Columbia, Vancouver, BC, Canada

keywords:
  - 

filters:
  - authors-block

notebook-links: false
number-sections: false
citation:
  container-title: Unpublished
# date: last-modified
bibliography: EIHurst.bib
csl: nature.csl

crossref:
  fig-title: '**Figure**'
  tbl-title: '**Table**'
  title-delim: "**.**"
---


```{r}
#| label: setup
#| include: false
# options
knitr::opts_chunk$set(
  # fig.width=8, fig.height=5,
  # out.width="50%",
  # fig.align="center",
  echo=FALSE,
  message=FALSE,
  warning=FALSE
  # cache=TRUE
)
set.seed(1234) # reproducible
options(knitr.kable.NA = '') # how kable handles NA
options(reticulate.repl.quiet = TRUE)
```

```{r}
#| label: libraries
#libraries
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
library(english) # numbers to words
library(tools) # title-case
capitalize_first <- function(x) {
  paste0(toupper(substr(x, 1, 1)), substr(x, 2, nchar(x)))
}
```

```{r}
load("_analysis/results.RData")
```

## Abstract

A 3T multimodal MRI study of healthy adults (n=19; 10 female; 21-54 years) was performed to investigate the potential link between fMRI long-range temporal correlations and excitatory/inhibitory balance. The study objective was to determine if the Hurst exponent (HE) – an estimate of the self-correlation and signal complexity of the blood-oxygen-level-dependent (BOLD) signal – is correlated with the excitatory-inhibitory (E/I) ratio. Findings in this domain have implications for neurological and neuropsychiatric conditions with disrupted E/I balance, such as autism spectrum disorder, schizophrenia, and Alzheimer’s disease. From a methodological perspective, HE is also considerably easier to accurately measure than E/I ratio. If HE can serve as a proxy for E/I, it may serve as a useful clinical biomarker for E/I imbalance. Moreover, E/I has been proposed to serve as a control parameter for brain criticality, which HE is believed to be a measure of. Thus, understanding if HE and E/I are correlated would serve to clarify this relationship. The study collected movie-watching and rest data including fMRI – which was used to calculate HE – and magnetic resonance spectroscopy (MRS) – which was used to measure inhibitory and excitatory neurotransmitters – GABA and glutamate, respectively. HE was found to increase with movie compared to rest, while E/I did not change between conditions. HE and E/I were not correlated during either movie or rest. This study represents the first attempt to investigate this connection *in vivo* in humans. We conclude that, at 3T and with our particular methodologies, no association was found.

# Introduction

Thirty years ago, functional magnetic resonance imaging (fMRI) profoundly changed the world of MRI by allowing real-time analysis of pressing neuropsychological questions [@ogawaMagneticResonanceImaging1990; @ogawaBrainMagneticResonance1990; @stephanShortHistoryCausal2012]. While initially used to probe task-based responses, researchers have more recently developed an interest in studying brain function at rest, known as resting-state fMRI (rs-fMRI) [@decoRestingBrainsNever2013], i.e. to understand how brain dynamics at rest are related to neurological functioning as well as individual differences. A critical tool in analyzing these dynamics is the Hurst exponent (HE) [@campbellMonofractalAnalysisFunctional2022], a measure of self-similarity derived from the blood-oxygen-dependent (BOLD) signal. HE estimates the extent to which the BOLD signal displays long-term memory, where a higher value indicates a self-similar signal with long-term positive autocorrelations [@campbellMonofractalAnalysisFunctional2022; @beggsBeingCriticalCriticality2012]. Another way of understanding HE is that a signal with high HE is fractal: similar temporal signal fluctuations are observed, no matter the time scale [@campbellMonofractalAnalysisFunctional2022].

HE has also emerged as a valuable tool in clinical research, capturing changes in BOLD signal dynamics across various neuropsychiatric conditions. In aging populations for instance, HE has been found to be elevated [@dongHurstExponentAnalysis2018; @winkAgeCholinergicEffects2006]; this relationship has also been found in mild cognitive impairment and Alzheimer’s disease [@maximFractionalGaussianNoise2005; @longBrainnetomeAtlasBased2018]. Additionally, changes in HE have been observed in conditions such as autism, mood disorders, and brain injury [@laiShiftRandomnessBrain2010; @donaTemporalFractalAnalysis2017; @weiIdentifyingMajorDepressive2013; @jingIdentifyingCurrentRemitted2017; @donaFractalAnalysisBrain2017]. These differences suggest HE may reflect changes in global and local functioning.

Underlying these observations is the critical brain hypothesis, which posits that the brain operates at a critical point, a state where order and disorder are perfectly balanced to enable optimal information processing [@decoRestingBrainsNever2013; @beggsBeingCriticalCriticality2012; @barangerChaosComplexityEntropy2000; @bassettUnderstandingComplexityHuman2011; @zimmernWhyBrainCriticality2020; @liangExcitationInhibitionBalance2024; @poilCriticalStateDynamicsAvalanches2012; @lombardiBalanceExcitationInhibition2017; @baumgartenCriticalExcitationinhibitionBalance2019; @bruiningMeasurementExcitationinhibitionRatio2020; @trakoshisIntrinsicExcitationinhibitionImbalance; @gaoInferringSynapticExcitation2017; @tianTheoreticalFoundationsStudying2022; @rubinovNeurobiologicallyRealisticDeterminants2011]. When operating at a critical point, the brain is maximally sensitive to external stimuli, and can dynamically transition between ordered and disordered states to promote efficient cognitive processing [@decoRestingBrainsNever2013; @beggsBeingCriticalCriticality2012; @tianTheoreticalFoundationsStudying2022; @rubinovNeurobiologicallyRealisticDeterminants2011].Recent papers suggest the control parameter underlying the brain’s ability to transition between states is the excitatory-inhibitory (E/I) ratio, the balance of excitatory and inhibitory neural activity, often operationalized as the ratio of the primary excitatory-to-inhibitory neurotransmitters, i.e. glutamate-to-GABA ratio [@liangExcitationInhibitionBalance2024; @lombardiBalanceExcitationInhibition2017; @baumgartenCriticalExcitationinhibitionBalance2019; @bruiningMeasurementExcitationinhibitionRatio2020; @trakoshisIntrinsicExcitationinhibitionImbalance; @gaoInferringSynapticExcitation2017]. It is thought that E/I controls criticality by modulating the brain’s signal-to-noise ratio [@liangExcitationInhibitionBalance2024; @rubensteinModelAutismIncreased2003].

Besides the implications to the critical brain hypothesis, understanding if E/I is related to HE may allow for easier estimation of excitatory and inhibitory neurotransmitters, as accurate E/I measurement is technically challenging [@ajramContribution1HMagnetic2019]. Magnetic resonance spectroscopy (MRS) is the only non-invasive method of measuring the ratio of glutamate (excitatory) to GABA (inhibitory) *in vivo* [@stanleyFunctionalMagneticResonance2018]. Unfortunately, it has both limited spatial and temporal resolution [@gaoInferringSynapticExcitation2017; @ajramContribution1HMagnetic2019; @stanleyFunctionalMagneticResonance2018]. Consequently, if HE could serve as a proxy for E/I, it would be much easier to estimate E/I in conditions of interest such as autism, Alzheimer’s, and schizophrenia.

There are a handful of studies suggesting a link between HE and E/I, however they are all either computational models or animal studies [@liangExcitationInhibitionBalance2024; @poilCriticalStateDynamicsAvalanches2012; @lombardiBalanceExcitationInhibition2017; @baumgartenCriticalExcitationinhibitionBalance2019; @bruiningMeasurementExcitationinhibitionRatio2020; @trakoshisIntrinsicExcitationinhibitionImbalance; @gaoInferringSynapticExcitation2017]. Moreover, their findings are inconsistent, with some reporting positive linear, negative linear, or U-shaped relationships between the two variables (see @tbl-lit). Thus, there is a need for further study, both to clarify the nature of a potential E/I-Hurst relationship, and also to confirm if this relationship indeed exists in the human brain. Therefore, the current study seeks to investigate the potential E/I-Hurst relationship in vivo, within the visual cortex during movie-watching and rest.

| Citation | Study Type | HE Data Type | HE Calculation Method | E/I Type | E/I-Hurst Relationship |
| ------------- | -------------- | -------------- | ------------- | -------------- | -------------- |
| Poil et al. (2012)@poilCriticalStateDynamicsAvalanches2012 | Computational with in-house simulated model | Neuronal avalanche size | Detrendend fluctuation analysis (DFA) | Structural: number of E-to-I neurons | Inverse U |
| Bruining et al. (2020)@bruiningMeasurementExcitationinhibitionRatio2020 | Computational with model by Poil et al. (2012); modified in-house | Neuronal oscillation amplitude | DFA | Structural: number of E-to-I synapses | Inverse U |
| Gao et al. (2017)@gaoInferringSynapticExcitation2017 | Computational; in vivo in rats and macaques | Local field potential (LFP) | PSD | Estimated from LFP | Positive linear |
| Lombardi et al. (2017)@lombardiBalanceExcitationInhibition2017 | Computational with in-house model | Neuronal avalanche size | PSD | Structural: number of E-to-I neurons | Negative linear |
| Trakoshis et al. (2020)@trakoshisIntrinsicExcitationinhibitionImbalance | Computational with simulated data; in vivo in mice | fMRI BOLD signal | Wavelet-based maximum likelihood method | E-to-I synaptic conductance | Positive linear |

: Summary of Methods for Existing E/I-Hurst Studies {#tbl-lit}


# Methods

## Study Participants

Ethics approval was granted by the Clinical Research Ethics Board at the University of British Columbia and BC Children's & Women's Hospital (H21-02686). Written informed consent was obtained from all participants. `r capitalize_first(as.character(english(as.numeric(count(orig_subjects)))))` participants were scanned between the ages of `r round(min(orig_subjects$Age),1)` and `r round(max(orig_subjects$Age),1)`.

After our analysis and performing quality assurance (see below), `r as.character(english(as.numeric(count(orig_subjects) - count(subjects))))` participants were removed for having poor quality data, leaving `r as.character(english(as.numeric(count(subjects))))` final participants, between the ages of `r round(min(subjects$Age),1)` and `r round(max(subjects$Age),1)` (mean ± sd: `r round(mean(subjects$Age),1)` ± `r round(sd(subjects$Age),1)` years; `r sum(subjects$Sex == "M")` males).

